---
title: "Problem Set 2"
author: "Patrick Altmeyer, Helena Patterson, Daniel Mueck and Gabriela Lavagna"
date: "`r format(Sys.Date(), '%d %B, %Y')`"
output: 
  bookdown::html_document2:
    code_folding: hide
    number_sections: false
    toc: true
    toc_float: true
  bookdown::pdf_document2:
    toc: false
    includes:
      in_header: preamble.tex
    number_sections: false
bibliography: bib.bib
---

**NOTE**: All code is hidden by default, but buttons can be used to unhide code chunks.

```{r setup, include=FALSE}
rm(list=ls())
knitr::opts_chunk$set(echo = TRUE, message=FALSE, warning = FALSE)
library(data.table)
```

```{r, eval=FALSE}
dt <- data.table(haven::read_dta("data/ps2.dta"))
fwrite(dt, "data/ps2.csv")
```

```{r}
dt <- fread("data/ps2.csv")
```


# 1.

Investigate bank balance-sheet data and show whether IRB- banks use internal models to reduce capital holdings. You should use both statistical and graphical tools.

<hr>

To make things clearer we reformat the date variable `date_q`:

```{r}
library(zoo)
dt[,date_q:=zoo::as.yearqtr(2000 + (date_q-min(date_q))/4)]
```

The table below shows a summary of the distribution of bank's total assets. Figure \@ref(fig:hist) provides a visual summary. The left panel shows the log-distribution, while the right panel shows levels in millions of assets. Light-blue bins indicate IRB-labelled banks.

```{r, results='asis'}
library(xtable)
options(xtable.comment = FALSE)
xtab = data.frame(c(summary(dt[date_q=="2007 Q1"]$bsize_)))
colnames(xtab) = "Total assets (logs)"
xtab <- xtable(xtab, label = "Distribution of banks' total assets as of 2017 Q1.")
print(xtab, type="html")
```

```{r hist, fig.width=10, fig.height=4, fig.cap="Histograms for banks' total assets as of 2017 Q1. The left panel shows the log-distribution, while the right panel shows levels in millions of assets. Light-blue bins indicate IRB-labelled banks."}
library(ggplot2)
theme_set(theme_bw())
dt[,total_assets_mn:=exp(bsize_)/1e6]
# Assign indicator:
dt_sub <- unique(dt[date_q=="2007 Q1", .(bankid, bsize_)])[,type:=ifelse(bsize_>quantile(bsize_, 0.75),"IRB", "SA")]
dt_sub[,bsize_:=NULL]
setkey(dt, bankid)
setkey(dt_sub, bankid)
dt <- dt_sub[dt]
# Prepare plot:
dt_plot <- melt(dt[date_q=="2007 Q1"], id.vars = c("firmid", "bankid", "date_q", "type"), measure.vars = c("bsize_", "total_assets_mn"))
levels(dt_plot$variable) <- c("Logs", "Levels (millions)")
p <- ggplot(data=dt_plot, aes(x=value, fill=type)) +
  geom_histogram(colour="blue") +
  scale_fill_manual(
    name="Risk-weighting:",
    values=c("lightblue", "white")
  ) +
  labs(
    x = "Total assets",
    y = "Count"
  ) +
  facet_wrap(
    ~ variable,
    scales = "free"
  )
p
```

The Basel III leverage ratio can be computed form the available data as $\ell_{\text{Basel}}= \frac{\ell_{\text{Tier-1}}A_{\text{risk-weighted}}}{A}$ where $\ell_{\text{Tier-1}}$ denotes the Tier-1 Capital Ratio, $A_{\text{risk-weighted}}$ denotes risk-weighted capital and $A$ denotes banks' total assets. Since assets (`bcet1_`,`bsize_`) are provided in log-terms we need to make a small adjustment for that and take exponentials (although not accounting for only leads to a relatively error):

```{r}
dt[,basel_lev:=(bcet1_*exp(bsizerw_))/exp(bsize_)] # compute leverage ratio
```

Then the following code generates Figure \@ref(fig:lev). The top panel shows time series of the different leverage ratios differentiated by bank type. The black vertical stalk indicates 2007 Q1. The bottom panel shows the difference in leverage ratios between IRB- and SA-labelled banks indexed to 2007 Q1. Upon visual inspection the difference in Basel III leverage ratios increased following the regulatory change in 2007 Q1. In the annex you may find a table of average leverage ratios pre- and post-treatment (over the period of analysis goes from 2006 Q1 to 2008 Q2).

```{r lev, fig.cap="Evolution of leverage ratio over time. The top panel shows time series of the different leverage ratios diffentiated by bank type. The black vertical stalk indicates 2007 Q1. The bottom panel shows the difference in leverage ratios between IRB- and SA-labelled banks indexed to 2007 Q1."}
dt_plot <- unique(melt(dt, id.vars = c("bankid", "date_q", "type"), measure.vars = c("bcet1_", "basel_lev")))
dt_plot <- dt_plot[,.(value=mean(value)),by=.(date_q, type, variable)][order(date_q, type, variable)] # averages
levels(dt_plot$variable) <- c("Tier-1", "Basel III")
p1 <- ggplot(data=dt_plot, aes(x=date_q, y=value, colour=type, linetype=variable)) +
  geom_line() +
  scale_color_discrete(name="Risk-weighting:") +
  scale_linetype_discrete(name="Ratio:") +
  labs(
    x="",
    y="Ratio"
  ) +
  geom_vline(xintercept = zoo::as.yearqtr("2007 Q1")) +
  theme(
    legend.position="top",
    axis.title.x = element_blank(),
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank()
  )
dt_tab <- copy(dt_plot[date_q %between% c("2006 Q1", "2008 Q2")]) # save for table in annex

dt_plot <- dt_plot[,.(value=value[type=="SA"] - value[type=="IRB"]), by=.(date_q, variable)]
dt_plot <- dt_plot[,.(date_q=date_q,value=value),by=variable]
p2 <- ggplot(dt_plot, aes(x=date_q, y=value, linetype=variable)) +
  geom_col(position = "dodge", fill="white", colour="black") +
  labs(
    x="Time",
    y="IBR minus SA"
  ) +
  scale_linetype_discrete(name="Ratio:") + 
  theme(legend.position="none") +
  scale_x_yearqtr(
    format = "%YQ%q"
  )

library(gridExtra)
grid.arrange(p1,p2, heights=c(2,1), ncol=1)
```

\FloatBarrier

# 2.

## Diff-in-diff

### Pooled OLS

Since the exercise clearly states that the period of analysis goes from 2006 Q1 to 2008 Q2, we restrict the sample for this exercise:

```{r}
dt_full <- copy(dt) # keep copy of full sample
dt <- dt[date_q %between% c("2006 Q1", "2008 Q2")]
```

Below we first run a difference-in-differences (DiD) model trough pooled OLS. We test a number of model specifications, but the coefficient of interest on `post:D` remains insignificant.

```{r, results='asis'}
library(stargazer)
dt[,post:=ifelse(date_q<="2007 Q1",0,1)] # define structural break indicator
dt[,D:=ifelse(type=="IRB",1,0)] # define treatment dummy
mod_pooled <- lm(lncredit ~ post * D, data = dt)
mod_pooled_X_macro <- lm(lncredit ~ post * D + mpshock + mprate + gdp + infl, data = dt)
mod_pooled_X_entity <- lm(lncredit ~ post * D + basel_lev + bsizerw_ + bsize_ + rating + empl + fsize, data = dt)
mod_pooled_X_all <- lm(lncredit ~ post * D + mpshock + mprate + gdp + infl + basel_lev + bsizerw_ + bsize_ + rating + empl + fsize, data = dt)
mods <- list(
  mod_pooled,
  mod_pooled_X_entity,
  mod_pooled_X_macro,
  mod_pooled_X_all
)
names(mods) <- c(
  "Simple",
  "Entity controls",
  "Macro controls",
  "Full controls"
)
cov_labels <- gsub("_","",attr(mod_pooled_X_all$terms,"term.labels"))
stargazer(
  mods,
  header = FALSE,
  type = "html",
  column.labels = names(mods),
  covariate.labels = cov_labels
)
```

### Including firm-bank fixed effects

In the table below we control for firm-bank fixed effects. We find a statistically significant increase in the difference in lending (logs) between IRB and SA banks.

```{r, results='asis'}
library(lfe)
dt[,firmbank:=paste(firmid,bankid,sep="_")]
mod_fe <- felm(lncredit ~ post * D | firmbank, data = dt)
mod_fe_X_macro <- felm(lncredit ~ post * D + mpshock + mprate + gdp + infl | firmbank, data = dt)
mod_fe_X_entity <- felm(lncredit ~ post * D + basel_lev + bsizerw_ + bsize_ + rating + empl + fsize | firmbank, data = dt)
mod_fe_X_all <- felm(lncredit ~ post * D + mpshock + mprate + gdp + infl + basel_lev + bsizerw_ + bsize_ + rating + empl + fsize | firmbank, data = dt)
mods <- list(
  mod_fe,
  mod_fe_X_entity,
  mod_fe_X_macro,
  mod_fe_X_all
)
names(mods) <- c(
  "Simple",
  "Entity controls",
  "Macro controls",
  "Full controls"
)
cov_labels <- gsub("_","",attr(mod_pooled_X_all$terms,"term.labels"))
stargazer(
  mods,
  header = FALSE,
  type = "html",
  column.labels = names(mods),
  covariate.labels = cov_labels
  # column.sep.width = "-5pt",
  # font.size = "footnotesize"
)
```


\FloatBarrier

## Parallel trend

```{r}
alpha <- .05
mod <- summary(felm(lncredit ~ factor(date_q) * D | firmbank, data = dt))
coef <- data.table(mod$coefficients)
coef[,name:=rownames(mod$coefficients)]
coef <- coef[name!="D"]
coef <- coef[grepl(":",name)] # keep only coeffs on interaction term
coef[,date:=dt[,unique(date_q)[-1]]]
setnames(coef, c("Estimate", "Std. Error"), c("estimate", "se"))
p <- ggplot(data = coef, aes(x=date, y=estimate)) +
  geom_errorbar(aes(ymin=estimate-se, ymax=estimate+se), width=.1) +
  geom_point() +
  geom_vline(xintercept = as.yearqtr("2007 Q1"), colour="red", linetype="dashed") +
  geom_hline(yintercept = 0) +
  labs(
    x="Time",
    y="Coefficient"
  ) + 
  scale_x_yearqtr(
    format = "%YQ%q"
  )
p
```


# Annex

```{r, results='asis'}
dt_tab[,post:=ifelse(date_q<="2007 Q1",0,1)]
dt_tab <- dt_tab[,.(value=mean(value)),by=.(type,variable,post)]
dt_tab <- dcast(dt_tab, type + variable ~ post, value.var = "value")
setnames(dt_tab,colnames(dt_tab),c("Bank type", "Leverage ratio", "Pre-treatment", "Post-treatment"))
xtab <- xtable(dt_tab)
print(xtab, type="html")
```



